\section{Project Plan}

\begin{enumerate}

\item Design and implement a programming language workbench for
experimenting with mobility-oriented programming contrasted with other
programming paradigms for distributed mobile cloud applications.

This is expected to result in scientific contributions in the field of
programming language design and implementation. Possible venues for
publication include <Programming>, ECOOP, POPL, PLDI, and ETAPS conferences.

This is in progress.

\item Design and implement a course on ``Programming Paradigms for
Distributed Mobile Cloud Programming''.

Focus on online content delivery and automated grading (i.e., build a
MOOC). This enables large samples of the target population to
participate, and complete the course. It also enables controlled
variability in what is taught to selected groups in the experiment.

This is expected to result in scientific contributions in terms of how
distributed programming exercises can be reliably graded
automatically. To this end, I plan to employ property-based
testing\cite{2000-QuickCheck, 2006-Telecoms-QuickCheck}, with adequate
extensions for stating properties about distributed (mobile cloud)
systems. Possible venues for publication include ACM Learning at
Scale, PLATEU.

This is in progress.

\item Run multiple iterations of the MOOC, report on the results, and
experiences with the research methodology. This too is expected to
result in scientific contributions. Possible venues for publication
include <Programming>, PLATEU, ECOOP, POPL, PLDI, and ETAPS
conferences.

\end{enumerate}

The current formal trajectory is to finish by November 30, 2021. If
the above work is completed by that time, most of the above may well
be complete, but will probably not yet be published. I have begun
investigating possibilities for formal extensions of this deadline.

% \begin{description}
% 
% \item[September 2020 -- November 2020]\hfill
% 
% \begin{itemize}
% 
% \item Seek feedback on the design of experiments for the evaluation of
% eloquence of programming models
% 
% \end{itemize}
% 
% Potential venues for paper submission:
% 
% \begin{itemize}
% 
% \footnotesize
% 
% \item PLATEAU 2021
% 
% \item European Joint Conferences on Theory \& Practice of Software
% (ETAPS) 2021
% 
% \item The Art, Science, and Engineering of Programming (<Programming>)
% 2021
 
% \item Programming Languages Design and Implementation (PLDI) 2021
% 
% \item International Conf. on Ambient Systems, Networks and
% Technologies (ANT) 2021
% 
% \item International Conf. on Emerging Data and Industry 4.0 (EDI40)
% 2021
% 
% \end{itemize}
% 
% \item[December 2020 -- February 2021]\hfill
% 
% Potential venues for paper submission:
% 
% \begin{itemize}
% 
% \footnotesize
% 
% \item European Conference on Object-Oriented Programming (ECOOP) 2021
% 
% \item USENIX Annual Technical Conference (USENIX ATC) 2021
% 
% \item Workshop on Modern Language Runtimes, Ecosystems, and VMs
% (MoreVMs) 2021
% 
% \item International Conference on Languages Compilers, Tools and
% Theory of Embedded Systems (LCTES) 2021
% 
% \end{itemize}
% 
% \item[March 2021 -- May 2021]\hfill
% 
% Potential venues for paper submission:
% 
% \begin{itemize}
% 
% \footnotesize
% 
% \item International Conference on Functional Programming (ICFP) 2021
% 
% \item International Conference on Types for Proofs and Programs
% (TYPES) 2021
% 
% \item International Conf. on Mobile Systems and Pervasive Computing
% (MobiSPC) 2020
% 
% \item Onward! 2021
% 
% \item CurryOn 2021
% 
% \item OOPSLA 2021
% 
% \item ICOOOLPS 2021
% 
% \item Erlang 2021
% 
% \item International Workshop on Context-Oriented Programming (COP)
% 2021
% 
% \end{itemize}
% 
% \item[June 2021 -- August 2021]\hfill
% 
% Potential venues for paper submission:
% 
% \begin{itemize}
% 
% \footnotesize
% 
% \item The Art, Science, and Engineering of Programming (<Programming>)
% 2022
% 
% \item International Conference on Software Language Engineering (SLE)
% 2021
% 
% \item PLATEAU 2022
% 
% \end{itemize}
% 
% \item[September 2021 -- November 2021]\hfill
% 
% \begin{itemize}
% 
% \item Write dissertation
% 
% \end{itemize}
% 
% \end{description}
